### 2013-06-08 doc/pgas13/evaluation.tex
Remove 'full' from plots, fix discusion, rename to 'dist' and 'centralized'

\paragraph{Queue and Stack}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/vector_perf.pdf}
  \caption{\emph{Queue and Stack performance.}
    Results are shown on a log scale for a throughput workload performing 256 million operations with 2048 workers per core and 16 cores per node. Flat combining improves throughput by at least an order of magnitude and allows performance to scale. Matching pushes and pops enables the stack to perform even better on a mixed workload.
  }
  \label{fig:vector}
\end{figure}

@@

Despite Grappa's automatic aggregation, without combining, both the stack and queue completely fail to scale because all workers' updates must serialize on the master core.
Though combining on the master core alleviates some of the serialization, it does not provide significant benefit because all of the synchronization must still be performed on one core. However, with local flat combining, synchronization is able to be done mostly in parallel, with much less-frequent bulk synchronization at the master. Doing the additional level of combining at the master seems to not affect performance significantly. This is somewhat surprising as all combined updates must completely serialize on the master still. However, with full combining enabled, we see that relatively few requests are combined together---for an all-push workload, on the order of a thousand are combined each pass, and for the mixed push/pop workload, on average less than 20 are combined.

@@

The queue also benefits from reduced synchronization and batching operations, and its all-push workload performs identically to the stack's.
However, the queue is unable to do matching locally, and in fact, the mixed workload performs worse because the current implementation serializes combined enqueue and dequeue operations. This restriction could be lifted with more careful synchronization at the master core allowing enqueues and dequeues to proceed in parallel as long as enough elements are in the queue to allow them to proceed without conflicting.

>>>>

\paragraph{Queue and Stack}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/vector_perf.pdf}
  \caption{\emph{Queue and Stack performance.}
    Results are shown on a log scale for a throughput workload performing 256 million operations with 2048 workers per core and 16 cores per node. Local flat combining improves throughput by at least an order of magnitude and allows performance to scale. Matching pushes and pops enables the stack to perform even better on a mixed workload.
  }
  \label{fig:vector}
\end{figure}

@@

Despite Grappa's automatic aggregation, without combining, both the stack and queue completely fail to scale because all workers' updates must serialize on the master core.
Though combining on the master core alleviates some of the serialization, it does not provide significant benefit because all of the synchronization must still be performed on one core. However, with local flat combining, synchronization is able to be done mostly in parallel, with much less-frequent bulk synchronization at the master.

@@

We also experimented with performing both forms of combining together; however, combining at the master did not affect performance significantly. This was somewhat surprising as combined updates will still force others to wait. However, with both levels of combining enabled, relatively few requests are combined at the master---for the all-push workload, on the order of a thousand, and for the mixed workload on the stack, the average is less than 20.

The queue benefits in the same way from reducing synchronization and batching, and its all-push workload performs identically to the stack's.
However, the queue is unable to do matching locally, and in fact, the mixed workload performs worse because the current implementation serializes combined pushes and combined pops. This restriction could be lifted with more careful synchronization at the master core allowing pushes and pops to proceed in parallel as long as there are enough elements in the queue.

### 2013-06-08 doc/pgas13/flat-combining.tex
Remove 'full' from plots, fix discusion, rename to 'dist' and 'centralized'

With flat combining, instead of all workers on a core sending independent messages to do remote synchronization, they can do combining locally and share a single remote synchronization operation. 
As with shared-memory combining, distributed combining improves locality, reduces the amount of global synchronization, and as bonus, allows for data structure-specific combining to magnify the benefits even further.

% allows synchronization to happen in parallel on all the cores, 
% Flat combining in Grappa is about allowing synchronization and  distributing synchronization 
% In a way, Grappa's automatic message aggregation is already trying to reduce synchronization traffic; however, the generic mechanism for serializing and packing has a cost, and in the end, the operations must still serialize on the master.
% Flat combining is about ``teaching'' it how to more efficiently perform many operations remotely in certain special cases.

@@

When a flat-combined global data structure is created, each core allocates a local proxy object in it's core-local heap. This proxy serves the role of the publication list, allowing workers within a core to coordinate their efforts. 
Conceptually, workers add their requests to the local publication list, and one is chosen to do the combining. However, because workers are scheduled cooperatively, each worker has atomic access to the local proxy ``for free'', so an explicit publication list is unnecessary.
Instead, each worker performs its own combining, and blocks if it was unable to satisfy its request immediately. Figure~\ref{fig:stackfc} shows how in the global stack example, pushes and pops can be matched locally, avoiding global synchronization for those operations.

@@

\section{Grappa FC Framework}
\TODO{add code examples, and add discussion of "master" combining}
% Many different data structures can integrate flat combining into their operation using the same delegation mechanism, making it easier to implement concurrent data structures.
In order to leverage the flat-combining paradigm in Grappa, we implemented a generic flat-combining framework which can be used to improve performance for a number of global data structures. The FC framework handles the common problems of managing the combiners, handling worker wake-ups, and maintaining progress. When hooking into the framework, each data structure implementation need only define how to combine operations and globally commit them.

% As discussed before, Grappa's distributed combining takes a different approach to expressing how operation locally  flat combining design takes a different approach to performing 
The Grappa FC framework takes a different approach than the original flat-combining work for expressing how operations combine. As mentioned before, the local ``publication list'' is trivial to access atomically in Grappa due to workers being cooperatively scheduled.
Instead, in the Grappa FC Framework, a \emph{proxy} object is instantiated ahead of time on each core and each worker merges its update into that structure before blocking or becoming the combiner.
% In the original work, an explicit publication list is built up, and when the combiner scans over the list, it has a structure that it uses to accumulate requests in a logical way. In the case of the stack, for instance, it is a thread-local stack, which it can use to immediately match up pushes and pops. In the Grappa implementation, each worker's actions are atomic anyway, so rather than creating an explicit request list, the combiner object (or \emph{proxy}) is instantiated ahead of time for the data structure, and each worker does its own local combining into that structure before blocking. This reduces the space blowout of a large dynamically-sized publication list (with associated caching issues), and does not overly affect the expression of combining.
In this style, each global data structure must define the following:
\begin{enumerate}
  \item A local \emph{proxy} data structure for tracking updates to the global structure.
  \item \emph{Combining methods} that operate on the local proxy.
  \item A \emph{sync} method that globally commits the proxy's state.
\end{enumerate}
An example proxy declaration for the GlobalStack is shown in Figure~\ref{fig:stack_proxy}.

@@

void push(GlobalAddress<GlobalStack> stack, T val) {
  // perform operation on the stack's master core
  delegate::call(stack.core(), [stack,val]{
    // force other operations to wait
    stack->lock.acquire();
    // do remote write to the top of the stack
    // (so other workers may interleave)
    delegate::write(stack->top, val);
    // bump the top pointer
    stack->top++;
    stack->lock.release();
  }); // blocks until response arrives
}
\end{lstlisting}

\caption{Snippet of code for pushing to a global stack (without flat combining).}
\label{fig:push}
\end{figure}

>>>>

\paragraph{Centralized Combining.}
A first thought might be to directly apply the idea of flat combining to the serialization at the master node. The worker that acquires the lock on the master can walk through the requests of all the other workers waiting to acquire the global lock and perform combining on them. In the case of the GlobalStack, this would mean taking all of the push requests, matching them with the pops, and sending messages back to the remote workers with the results. Any leftover pushes or pops would be applied to the data structure, and another round of combining would begin. This approach, combining on the master, reduces traffic to the data structure storage, but the single ``master'' core must still process every single request. This approach will not scale when every other core can generate requests just as fast as the master can combine them.

\paragraph{Distributed Combining}
Instead of all workers on a core sending independent synchronization messages and putting the burden all on the master, what if each core did its own combining first, and synchronized with the master in bulk?
Distributing the combining to each core allows the majority of the work to be performed in parallel and without communication.
As with shared-memory combining, distributed combining improves locality, reduces the amount of global synchronization, and allows operations to be composed in data structure-specific ways for even greater performance.
% allows synchronization to happen in parallel on all the cores, 
% Flat combining in Grappa is about allowing synchronization and  distributing synchronization 
% In a way, Grappa's automatic message aggregation is already trying to reduce synchronization traffic; however, the generic mechanism for serializing and packing has a cost, and in the end, the operations must still serialize on the master.
% Flat combining is about ``teaching'' it how to more efficiently perform many operations remotely in certain special cases.

@@

In distributed flat-combining, each core needs its own publication list to track all of its outstanding operations to be performed on the global data structure. In Grappa, this takes the form of a local \emph{proxy} object allocated from the core-local heap.
Conceptually, workers add their requests to the local publication list, one is chosen to do the combining, and the rest block until their request is satisfied.
However, because workers are scheduled cooperatively, each worker has atomic access to the local proxy ``for free'', so an explicit publication list is unnecessary.
Instead, each worker performs its own combining, pushing the 
, and blocks if it was unable to satisfy its request immediately. Figure~\ref{fig:stackfc} shows how in the global stack example, pushes and pops can be matched locally, avoiding global synchronization for those operations.

@@

\section{Grappa FC Framework}
% One of the strengths of flat combining is that the same combining mechanism can be applied in multiple data structures. Each structure must only specify how its operations compose.
% Many different data structures can integrate flat combining into their operation using the same delegation mechanism, making it easier to implement concurrent data structures.
In order to leverage the flat-combining paradigm in Grappa, we implemented a generic framework which can be used to improve performance for a number of global data structures. The FC framework handles the common problems of managing the combiners, handling worker wake-ups, and maintaining progress. When hooking into the framework, each data structure implementation need only define how to combine operations and globally commit them.

% As discussed before, Grappa's distributed combining takes a different approach to expressing how operation locally  flat combining design takes a different approach to performing 
The Grappa FC framework takes a different approach than the original flat-combining work for expressing how operations combine. As mentioned before, the local ``publication list'' is trivial to access atomically in Grappa due to workers being cooperatively scheduled.
Instead, in the Grappa FC Framework, a \emph{proxy} object is instantiated ahead of time on each core and each worker merges its update into that structure before blocking or becoming the combiner.
% In the original work, an explicit publication list is built up, and when the combiner scans over the list, it has a structure that it uses to accumulate requests in a logical way. In the case of the stack, for instance, it is a thread-local stack, which it can use to immediately match up pushes and pops. In the Grappa implementation, each worker's actions are atomic anyway, so rather than creating an explicit request list, the combiner object (or \emph{proxy}) is instantiated ahead of time for the data structure, and each worker does its own local combining into that structure before blocking. This reduces the space blowout of a large dynamically-sized publication list (with associated caching issues), and does not overly affect the expression of combining.
In this style, each global data structure must define the following:
\begin{enumerate}
  \item A local \emph{proxy} data structure for tracking updates to the global structure.
  \item \emph{Combining methods} that operate on the local proxy.
  \item A \emph{sync} method that globally commits the proxy's state.
\end{enumerate}
An example proxy declaration for the GlobalStack is shown in Figure~\ref{fig:proxy}.

@@

void push(GlobalAddress<GlobalStack> stack, T e) {
  // perform operation on the stack's master core
  delegate::call(stack.core(), [stack,e]{
    // force other operations to wait
    stack->lock.acquire();
    // do remote write to the top of the stack
    // (so other workers may interleave)
    delegate::write(stack->top, e);
    // bump the top pointer
    stack->top++;
    stack->lock.release();
  }); // blocks until response arrives
}\end{lstlisting}
\caption{Snippet of code for pushing to a global stack (without flat combining).}
\label{fig:push}
\end{figure}
